---
title: "model"
author: "poorva seth"
date: "16/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("Pmisc", repos='http://r-forge.r-project.org')
library(haven)
library(tidyverse)
library(dplyr)
library(lmtest)
library(ggplot2)
library(Pmisc)
nsse_data <- read_dta("/Users/evaseth/Desktop/NSSE.dta")
nsse_data$logdate <- NULL
nsse_data_2014 <- subset(nsse_data, nsse_data$Year == '2014')
```


```{r ONT01 obstacles}
nsse_data_2014$ONT01c <- ifelse(nsse_data_2014$ONT01c == 1, 0, 1)
nsse_data_2014$ONT01d <- ifelse(nsse_data_2014$ONT01d == 1, 0, 1)
nsse_data_2014$ONT01e <- ifelse(nsse_data_2014$ONT01e == 1, 0, 1)
nsse_data_2014$ONT01g <- ifelse(nsse_data_2014$ONT01g == 1, 0, 1)
nsse_data_2014$ONT01h <- ifelse(nsse_data_2014$ONT01h == 1, 0, 1)
```

```{r rename}
#ONT01g is at index 221, now named languagebarrier
names(nsse_data_2014)[221] <- "LanguageBarrier"

#ONT01h is index 222, now disabilitybarrier
names(nsse_data_2014)[222] <- "DisabilityBarrier"

#ONT01c is index 217
names(nsse_data_2014)[217] <- "AcadPerformanceBarrier"

#ONT01d is index 218
names(nsse_data_2014)[218] <- "ScheduleBarrier"

#ONT01e is index 219
names(nsse_data_2014)[219] <- "AdvisingBarrier"

#ONT05k is index 265
names(nsse_data_2014)[265] <- "NoCommunitySense"

#ONT04a is index 242
names(nsse_data_2014)[242] <- "MoreProfContact"

#ONT06h is index 282
names(nsse_data_2014)[282] <- "AcademicProbsInfo"

#ONT07 is index 290
names(nsse_data_2014)[290] <- "Residency"


#add internatYN var
nsse_data_2014$internatYN <- ifelse(nsse_data_2014$internat_ca == '3', 'Canadian Citizen', 'International')

# add internship var 
nsse_data_2014$internship <- ifelse(!nsse_data_2014$intern == "4", "0", "1")
# It's a 0 if they have not done it yet

#add combined qi feedback
nsse_data_2014$combined_feedback <- rowSums(nsse_data_2014[,c(30,31)], na.rm = TRUE)

#group faculties
nsse_data_2014$ARTSC <- ifelse(nsse_data_2014$group1 == "ARTSC", "1", "0")
nsse_data_2014$ERIN <- ifelse(nsse_data_2014$group1 == "ERIN", "1", "0")
nsse_data_2014$SCAR <- ifelse(nsse_data_2014$group1 == "SCAR", "1", "0")
nsse_data_2014$APSC <- ifelse(nsse_data_2014$group1 == "APSC", "1", "0")
nsse_data_2014$Other_Faculty <- ifelse(nsse_data_2014$group1 == "ARCLA" |nsse_data_2014$group1 == "FPEH" | nsse_data_2014$group1 == "MUSIC" | nsse_data_2014$group1 == "NURS", "1", "0")

#rename IRsex where being male means 1
nsse_data_2014$Male <- ifelse(nsse_data_2014$IRsex == "1", "1", "0" )

#rename IRClass to use as residency variable, 0 means no residence (fourth year), 1 means resdience (first year)
#nsse_data$residency <- ifelse(nsse_data$IRclass == "4", "0", "1" )

nsse_data_2014_first <- subset(nsse_data_2014, nsse_data_2014$IRclass == '1')
nsse_data_2014_fourth <- subset(nsse_data_2014, nsse_data_2014$IRclass == '4')
```


INPUTS: IRclass, LanguageBarrier, DisabilityBarrier, tmworkoffhrs, tmcarehrs, tmcommutehrs, re_white_ca, re_chin_ca, re_s_asia_ca, firstgen, begincol [0 means you started here], IRSex (0 meaning Female), ARTSC, APSC, ERIN, SCAR,internatYN

#no sexual orientation in 2014 data

ENVIRONMENTS: AcadPerformanceBarrier, ScheduleBarrier, AdvisingBarrier, NoCommunitySense, MoreProfContact, AcademicProbsInfo, tmworkonhrs, internship (would not let me separate by first or fourth years, but have done tables to show them in separate groups), HIP, athlete, combined_feedback, QI


#build a model
```{r model}
sfmodel_2014_first <- lm(SF ~  Residency + LanguageBarrier + DisabilityBarrier + tmworkoffhrs + tmcarehrs + tmcommutehrs + re_white_ca + re_chin_ca + re_s_asia_ca +  firstgen + internatYN + begincol + Male + ARTSC + ERIN + SCAR + APSC + AcadPerformanceBarrier + ScheduleBarrier + AdvisingBarrier + NoCommunitySense + MoreProfContact + AcademicProbsInfo + tmworkonhrs+ HIP + athlete + combined_feedback + QI , data = nsse_data_2014_first)

summary(sfmodel_2014_first)
```

```{r fourth year model}
sfmodel_2014_fourth <- lm(SF ~  Residency + LanguageBarrier + DisabilityBarrier + tmworkoffhrs + tmcarehrs + tmcommutehrs + re_white_ca + re_chin_ca + re_s_asia_ca +  firstgen + internatYN + begincol + Male + ARTSC + ERIN + SCAR + APSC + AcadPerformanceBarrier + ScheduleBarrier + AdvisingBarrier + NoCommunitySense + MoreProfContact + AcademicProbsInfo + tmworkonhrs+ HIP + athlete + combined_feedback + QI , data = nsse_data_2014_fourth)

summary(sfmodel_2014_fourth)
```

#working with 2017 data
```{r open data}
nsse_data <- read_dta("/Users/evaseth/Desktop/NSSE.dta")
nsse_data$logdate <- NULL
nsse_data_2017 <- subset(nsse_data, nsse_data$Year == '2017')

nsse_data_2017$ONT01c <- ifelse(nsse_data_2017$ONT01c == 1, 0, 1)
nsse_data_2017$ONT01d <- ifelse(nsse_data_2017$ONT01d == 1, 0, 1)
nsse_data_2017$ONT01e <- ifelse(nsse_data_2017$ONT01e == 1, 0, 1)
nsse_data_2017$ONT01g <- ifelse(nsse_data_2017$ONT01g == 1, 0, 1)
nsse_data_2017$ONT01h <- ifelse(nsse_data_2017$ONT01h == 1, 0, 1)

#relabel sexorient where straight is 0 and other is 1
nsse_data_2017$sexorient <- ifelse(nsse_data_2017$sexorient17 == 1, 0, 1)

#ONT01g is at index 221, now named languagebarrier
names(nsse_data_2017)[221] <- "LanguageBarrier"

#ONT01h is index 222, now disabilitybarrier
names(nsse_data_2017)[222] <- "DisabilityBarrier"

#ONT01c is index 217
names(nsse_data_2017)[217] <- "AcadPerformanceBarrier"

#ONT01d is index 218
names(nsse_data_2017)[218] <- "ScheduleBarrier"

#ONT01e is index 219
names(nsse_data_2017)[219] <- "AdvisingBarrier"

#ONT05k is index 265
names(nsse_data_2017)[265] <- "NoCommunitySense"

#ONT04a is index 242
names(nsse_data_2017)[242] <- "MoreProfContact"

#ONT06h is index 282
names(nsse_data_2017)[282] <- "AcademicProbsInfo"

#ONT07 is index 290
names(nsse_data_2017)[290] <- "Residency"


#add internatYN var
nsse_data_2017$internatYN <- ifelse(nsse_data_2017$internat_ca == '3', 'Canadian Citizen', 'International')

# add internship var 
nsse_data_2017$internship <- ifelse(!nsse_data_2017$intern == "4", "0", "1")
# It's a 0 if they have not done it yet

#add combined qi feedback
nsse_data_2017$combined_feedback <- rowSums(nsse_data_2017[,c(30,31)], na.rm = TRUE)

#group faculties
nsse_data_2017$ARTSC <- ifelse(nsse_data_2017$group1 == "ARTSC", "1", "0")
nsse_data_2017$ERIN <- ifelse(nsse_data_2017$group1 == "ERIN", "1", "0")
nsse_data_2017$SCAR <- ifelse(nsse_data_2017$group1 == "SCAR", "1", "0")
nsse_data_2017$APSC <- ifelse(nsse_data_2017$group1 == "APSC", "1", "0")
nsse_data_2017$Other_Faculty <- ifelse(nsse_data_2017$group1 == "ARCLA" |nsse_data_2017$group1 == "FPEH" | nsse_data_2017$group1 == "MUSIC" | nsse_data_2017$group1 == "NURS", "1", "0")

#rename IRsex where being male means 1
nsse_data_2017$Male <- ifelse(nsse_data_2017$IRsex == "1", "1", "0" )

#rename IRClass to use as residency variable, 0 means no residence (fourth year), 1 means resdience (first year)
#nsse_data$residency <- ifelse(nsse_data$IRclass == "4", "0", "1" )

nsse_data_2017_first <- subset(nsse_data_2017, nsse_data_2017$IRclass == '1')
nsse_data_2017_fourth <- subset(nsse_data_2017, nsse_data_2017$IRclass == '4')

```

```{r first year 2017model}
sfmodel_2017_first <- lm(SF ~   Residency + LanguageBarrier + DisabilityBarrier + tmworkoffhrs + tmcarehrs + tmcommutehrs + re_white_ca + re_chin_ca + re_s_asia_ca +  firstgen + internatYN + begincol + Male + ARTSC + ERIN + SCAR + APSC + AcadPerformanceBarrier + ScheduleBarrier + AdvisingBarrier + NoCommunitySense + MoreProfContact + AcademicProbsInfo + tmworkonhrs+ HIP + athlete + combined_feedback + QI , data = nsse_data_2017_first)

summary(sfmodel_2017_first)
```


```{r fourth year 2017}
sfmodel_2017_fourth <- lm(SF ~  Residency + LanguageBarrier + DisabilityBarrier + tmworkoffhrs + tmcarehrs + tmcommutehrs + re_white_ca + re_chin_ca + re_s_asia_ca +  firstgen + internatYN + begincol + Male + ARTSC + ERIN + SCAR + APSC + AcadPerformanceBarrier + ScheduleBarrier + AdvisingBarrier + NoCommunitySense + MoreProfContact + AcademicProbsInfo + tmworkonhrs+ HIP + athlete + combined_feedback + QI , data = nsse_data_2017_fourth)

summary(sfmodel_2017_fourth)
```

```