---
title: "STA442 Assignment 1"
author: "Poorva Seth"
date: "28/09/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(lmtest)
Affairs <- read_csv("Affairs (1).csv")
load("~/Desktop/smoke.RData")
```


# Question 1: Affairs. Use the Affairs dataset to address the research question/
# if having children has an effect on the chances of men and women having extramarital affairs.

### Report explaining my methods:

I used the 'Affairs' dataset which collected data from 600 married readers. The question of interest
concerned whether having children had an effect on the chances that men and women would have affairs.

I used a generalized linear model called `affairsmodel`, with a Bernoulli distribution. Each response 
follows a Binomial distribution, since the response variable is whether or not they have ever 
had an affair. The model can be described as follows:

$$ Y_i - Bernoulli(\mu_i) $$  
$$log(\mu_i/1-\mu_i) = X_i\beta$$  

$Y_i$: This is whether person i has ever had an affair. It follows a Bernoulli distribution 
for each individual and it linked to a linear combination of covariates using a logit link.  


$\mu_i$: This is the probability that the ith person has ever had an affair.  

$X_i$: This is a linear combination of covariates, consisting of `Children`, if the person has children,
`aged` which is the Age variable centred at 32 for easy interpretation, `Gender` of person i, `Yearsmarried` 
which is how long person i has been married, and also `religious`, which is a factored variable determining 
how religious person i is. There is also an interaction term between gender and children.

$\beta$: Is the vector of coefficients.  


The `Table 1 Affairs Model Summary` below shows the exponentiated log odds output, which was done to be able to interpret 
the results. In this case, the intercept would refer to a man who is 32 years old with children, has just been married 
and is not religious. The 95% confidence intervals in the table give us an idea of what covariates are considered to 
have a significant effect on ever having an affair. Some significant ones include `yearsmarried`, this is significant since 
the confidence interval does not include 1. The coefficient tells us that for every yearly increase of having been married, 
the odds are increased by 11%. Another significant factor is `religiousanti`, the confidence interval does not contain 1. 
The large coefficient indicates that being anti-religious increases your odds of having an affair by more than twice the 
baseline amount. Other levels of religiousness, like `religiouslow` and `religiousmed` were not deemed signficant by this model. 

`Table 2 and 3` show the odds ratios for the interaction terms between children and gender. The estimate from `Table 3` 
shows that men without children have decreased odds of having an affair by 22% compared to men with children. `Table 2`
shows that women without children have decreased odds of having an affair by 48% compared to women with children.

However, The confidence intervals in both tables contain 1, meaning they are not significant. We don't have 
enough evidence to conclude that sex and having children are signifcant predictors of whether or not one has an affair. 
This is not consistent with the research hypothesis, i.e that the odds of ever having an affair are influenced by your 
sex and whether or not you have children.

```{r q1, echo=FALSE}
Affairs$ever = Affairs$affairs > 0
Affairs$religious = factor(Affairs$religiousness,
                           levels=c(2,1,3,4,5), labels=c('no','anti','low','med','high'))
Affairs$aged = Affairs$age - 32

affairsmodel = (glm(ever ~ children:gender + aged + yearsmarried + religious, data=Affairs, family = 'binomial' ))
                           

logoddsaffairs = cbind(est=affairsmodel$coef, confint(affairsmodel, level = 0.95))


oddsaffairs = exp(logoddsaffairs)
oddsaffairs[1, ] = oddsaffairs[1, ]/(1 + oddsaffairs[1, ])
rownames(oddsaffairs)[1] = "Baseline Prob"
knitr::kable(oddsaffairs, digits = 3, format = "pandoc", caption="Affairs Model Summary") 


coefss= affairsmodel$coef[1:10]
varss= vcov(affairsmodel)[1:10,1:10]
FemalenoKids=c(1,0,0,0,0,0,0,1,0,0)
MalenoKids=c(1,0,0,0,0,0,0,0,0,1)
FemaleyesKids=c(1,0,0,0,0,0,0,0,1,0)
MaleyesKids=c(1,0,0,0,0,0,0,0,0,0)
FemaleRatio=FemalenoKids-FemaleyesKids
femaless <- t(FemaleRatio) %*% coefss
estims_female<-drop(femaless) 
se_female <- sqrt(drop(t(FemaleRatio) %*% varss %*% FemaleRatio))
final_females <- exp(estims_female+c(Estimate=0,Lower=-2,Upper=2)*se_female)
knitr::kable(cbind(final_females), digits = 2, format = "pandoc", caption="Odds for women without and with children")
MaleRatio=MalenoKids-MaleyesKids
maless <- t(MaleRatio) %*% coefss
estims_male<-drop(maless) 
se_male <- sqrt(drop(t(MaleRatio) %*% varss %*% MaleRatio))
final_males <- exp(estims_male+c(Estimate=0,Lower=-2,Upper=2)*se_male)
knitr::kable(cbind(final_males), digits = 2, format = "pandoc", caption="Odds for men without and with children")

```


### Conclusion:

Recent findings from a study at the University of Toronto may be a surprise to many. The study used data from 600 married readers 
to study if an individual's sex and whether they have children were significant influences on ever having had an affair. It is often 
perpetuated by the media and stereotypes that becoming a mother makes a woman less likely to cheat, while becoming a father drives a 
man to be more likely. Contrary to popular belief, the findings showed that sex and having children were *not* significant factors on 
the odds of having an affair! Whereas other variables that are often overlooked, such as how long one has been married for and how 
religious they are, *did* have a significant impact on the odds of having an affair as revealed by the study! This is an interesting 
example of how common conceptions about society may be based on fallacious reasoning.


# Question 2: Use the smoking dataset to address the research hypotheses:
# i) White Americans use cigars, cigarillos, or little cigars roughly the same amount as Hispanic or Black Americans, 
# accounting for the fact that White Americans are more likely to live in rural areas, where smoking is a common phenomenon
# ii) The chances of using an e-cigarette is the same for both sexes, when holding other demographic characteristics similar.


## Short Summary:
Data was collected from the 2019 American National Youth Tobacco Survey to answer the research hypotheses. It is hypothesized that White 
Americans use cigars, cigarillos, or little cigars roughly the same amount as Hispanic or Black Americans, accounting for the fact that 
White Americans are more likely to live in rural areas, where smoking is a common phenomenon. My findings from my model, which accounts 
for the kind of area they live in, fail to provide any evidence that the smoking use of Hipanic Americans differ from those of Whites. This
is consistent with the part of the research hypothesis concerning Hispanic Americans. My findings also provide evidence to show that smoking
is much more prevalent in Rural areas as opposed to Urban, which is consistent with the research hypothesis. However, it was also hypothesized 
that smoking is no more common with Whites than it is for Blacks. My model provides sufficient evidence to conclude that there is in fact a s
ignficant difference between the smoking usage between Blacks and Whites, and that use of cigars, cigarillos or little cigars is more common for 
Blacks by roughly 53.5%. This is inconsistent with the first research hypothesis. The second hypothesis says that the likelihood of having used 
an e-cigarette is the same for both sexes, holding other characteristics similar. After testing for the likelihood, my findings do not provide 
any evidence that sex is a significant predictor on the use of e-cigarettes while accounting for other demographic characteristics. This is 
consistent with the hypothesis. 


## Write Up:

### Introduction:
We were told to investigate the hypotheses that smoking of cigars, cigarillos, or little cigars is just as common amongst European Americans than 
it is amongst Hispanics and Blacks, and also that the likelihood of having used an e-cigarette is the same for both sexes, holding other
characteristics constant. This data came from the 2019 American National Youth Tobacco Survey. For the hypothesis concerning cigars 
and cigarillos, we had to account for the fact that Whites are more likely to live in rural areas, where smoking is more prevalent. 

### Methods:
For the first hypothesis, my method was to use a generalised linear model. The response variable was `ever_cigars_cigarillos_or`, 
which asked the respondent if they had ever tried a cigar, cigarillo, or little cigar. Since this response only had two answers,
'yes' or 'no', I used a binomial family for the GLM. The predictors needed to address this question were `Race`, `RuralUrban`, `Sex`, 
and `Age`, since these were important demographic characteristics which would allow us to fairly compare the response variable amongst 
ethnicities. Using a binomial family for my generalised linear model by default involves using a logit link. To be able to interpret 
the output, I exponentiated the coefficients for easy interpretation to get it back on the natural scale. Evaluating the confidence 
interval of each predictor allows us to see which ones significantly differ from the intercept. In this case, the intercept refers to 
White male Americans who live in Urban areas and are 16 years old. The model can be described as:

$$ Y_i - Bernoulli(\mu_i) $$  
$$log(\mu_i/1-\mu_i) = X_i\beta$$  

$Y_i$: This is whether person i has ever tried smoking a cigar, cigarillo or little cigar. It follows a bernoulli distribution 
for each individual and it linked to a linear combination of covariates using a logit link.  

$\mu_i$: This is the probability that the ith person has ever tried smoking a cigar, cigarillo, or little cigar.  

$X_i$: This is a linear combination of covariates, consisting of `Sex`, `aged`, `Race`, and `RuralUrban`. The `aged` variable is 
the original `Age` variable from the dataset, but has been centered around 16 year olds.  

$\beta$: Is the vector of coefficients.  



The output is shown below:  


```{r question 2a, echo=FALSE}

smokeSub = smoke[which(smoke$Age >= 10), ]
smokeSub$aged <- smokeSub$Age - 16

model1 <- (glm(ever_cigars_cigarillos_or ~ Sex + Race + aged + RuralUrban,family='binomial',data=smokeSub))

logoddssmokers = cbind(est=model1$coef, confint(model1, level = 0.95))


oddssmokers = exp(logoddssmokers)
oddssmokers[1, ] = oddssmokers[1, ]/(1 + oddssmokers[1, ])
rownames(oddssmokers)[1] = "Baseline Prob"
knitr::kable(cbind(oddssmokers), digits = 3, format = "pandoc", caption = "Smokers Model Summary")


```

For the second hypothesis, my method was to use a likelihood ratio test since we are comparing two nested models, 
only differing by adding `Sex` as a predictor. Both models were generalised linear models. The response variable was
`ever_cigars_cigarillos_or`, which asked the respondent if they had ever tried a cigar, cigarillo, or little cigar. 
Since this response only had two answers, 'yes' or 'no', I used a binomial family for the GLM and had to exponentiate
the results for interpretation, as last time. This test assesses the ratio of their likelihoods. In this case, the research
hypothesis is that there is no signficant difference between both models, meaning that `Sex` is not a signficant predictor 
if all other predictors are constant. Both models included other important demographic characteristics such as `Race`, `Age`, 
and `RuralUrban` because these are important confounders, and one model will have `Sex` since that is what we are testing for.
Evaluating the p value of the likelihood ratio test gives us evidence to either reject, or fail to reject the research hypothesis. 
My findings indicate that we should fail to reject the research hypothesis. 

The model can be described the same way as above, except now only one model includes `Sex` in the linear combination of covariates.  

The output is shown below.  

```{r question 2b, echo=FALSE}
smokeSub = smoke[which(smoke$Age >= 10), ]
smallersmokesub = na.omit(smokeSub[, c("Sex", "Age", "Race", "RuralUrban", "ever_ecigarette")])
smallersmokesub$aged <- smallersmokesub$Age - 16

model1 <- (glm(ever_ecigarette ~ Sex + Race + aged + RuralUrban,family='binomial', data=smallersmokesub))

model2 <- (glm(ever_ecigarette ~  Race + aged + RuralUrban,family='binomial', data=smallersmokesub))

knitr::kable(lrtest(model1, model2), digits = 3, format = "pandoc", caption = "Likelihood Test Results")

```


### Results:
Output from the  `Table 4: Smokers Model Summary`  allowed me to interpret results for the first hypothesis.The baseline 
refers to White urban male Americans who are 16 years old. When looking at the 95% confidence interval for `Racehispanic`, we can 
see that it includes 1. This indicates that there is no evidence to suggest that cigar or cigarello usage amongst Hispanics differs 
signficantly from Whites. This is consistent with the research hypothesis that smoking cigars is no more common amongst White Americans
than with Hispanic Americans. I included `RuralUrban` in my model to account for the different areas amongst individuals. We can see 
that the confidence interval for `RuralUrbanRural` does not include 1, indicating that living in a Rural area does have a significant 
positive affect on the use of cigars or cigarillos by roughly 49% as indicated by the coefficient. This is consistent with the information 
provided in the hypothesis that smoking is a rural phenomenon. The first hypothesis suggested that smoking is no more common with Whites than 
it is for Blacks. When looking at the 95% confidence interval for `Raceblack`, it does not contain 1. This suggests that being Black
significantly impacts use of cigars or cigarillos by an increase of roughly 53.5%, as indicated by the coefficient. This is not consistent
with the research hypothesis, so we do not have evidence to include that the first hypothesis is overall plausible.  The results from  `Table 5: 
Likelihood Test Results`  allows me to compare the two nested GLM models. The models differ only by the inclusion of `Sex`. It is important to
note that I am using an alpha level of .05 to determine signficance. The output shows a p value of .06, which is outside of our level of
significance. This p-value indicates there is not enough evidence to suggest that the inclusion of `Sex` as a covariate had any effect, so we
cannot include that `Sex` is significant. This is consistent with the research hypothesis that the likelihood of having ever used an 
electronic cigarette is the same between the two sexes, holding other characteristics to be similar. 
 


